---
title: "Productivity"
author: "Merle Nye"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(infer)
library(survminer)
library(survival)
library(hablar)
library(dplyr)
library(rms)
```

```{r data}
library(readr)
EIA_7A_Compiled2 <- read_csv("EIA-7A_Compiled_fixed.csv")
data <- EIA_7A_Compiled2
head(EIA_7A_Compiled2)
```

```{r filter}
data <- data%>%
  filter(mine.status == "Active")%>%
  filter(year >= 2001)%>%
  filter(!is.na(production.stons))%>%
  filter(!is.na(labor.hours))%>%
  filter(production.stons != 0)%>%
  mutate(productivity = (production.stons/ labor.hours))
  
```


```{r total}
total <- data%>%
  group_by(year)%>%
  summarize(bigt = sum(production.stons))
total
```

```{r Appalachia}
data <- data%>%
  mutate(appalachia = case_when(mine.state == "Kentucky (East)" ~ "Yes", mine.state == "Kentucky (West)" ~ "Yes", mine.state == "Tennessee" ~ "Yes", mine.state == "West Virginia" ~ "Yes", mine.state == "Virginia" ~ "Yes", mine.state == "West Virginia (Southern)" ~ "Yes", mine.state == "West Virginia (Northern" ~ "Yes" ))
```


Below, the mines are categorizd as either low productivity or high productivity over the course of the years in question. The median is determined by the median productivity from 2008-present.



```{r survival_function}
year_closed <- function(id){
  yearin = 2008
  downed <- 2016
  while(yearin <= 2016){
  current_year <- data%>%
    filter(year==yearin)
  current_year <- current_year[,'msha.id']
  if ((id %in% current_year$msha.id) == TRUE){
    yearin <- yearin+1
  }else{
    downed <- yearin - 1
    break
  }
  }
  return(downed)
}
 
```

```{r open_function}
year_open <- function(varia){
  yearer <- data_for_open%>%
    filter(rank == varia)%>%
    pull(year)
  id <- data_for_open%>%
    filter(rank == varia)%>%
    pull(msha.id)
  id <- id[1]
  prev_year <- data%>%
    filter(year==yearer-1)
  prev_year <- prev_year[,'msha.id']
  if ((id %in% prev_year$msha.id) == TRUE){
    return(1)
  }else{
    return(0)
  }
}
```

```{r apply_open_function}
data_for_open <- data%>%
  filter(year >= 2008)
 
data_for_open <- data_for_open%>%
  mutate(rank = 1:nrow(data_for_open))%>%
   mutate(newvar = (rank))
data_for_open%>%
  group_by((newvar))%>%
  summarize(count=n())
open_table <- 0
open_table <- lapply(data_for_open$newvar, year_open)

new_open_table <- as.data.frame(do.call(rbind,open_table))
new_open_table <- new_open_table%>%
  mutate(rank=1:nrow(new_open_table))

complete_open <- full_join(data_for_open, new_open_table)
```

For some reason the summarize function doesn't work. It could be caused by NAs, but anyway, I didn't need to run it again. For future condiseration, the data set used here includes all mines since 2008 while grand ONLY contains observations from that time that were ALSO in 2008
```{r open_effeciency}
new_vs_old_exclusive <- grand%>%
  group_by(year.x, V1)%>%
  summarize(effeciency = mean(productivity.x))
new_vs_old_inclusive <- new_vs_old_exclusive%>%
  group_by(year)%>%
  summarize(effeciency_total = mean(effeciency))
total_old_new <- left_join(new_vs_old_exclusive, new_vs_old_inclusive, by="year")

new_vs_old_avg <- complete_open%>%
  group_by(V1)%>%
  summarize(avg_eff = mean(productivity))
write.csv(total_old_new, file = "New_vs_Old_mine_effeciency")

```

```{r open_graphs}
ggplot(new_vs_old_exclusive, aes(x=year, y =effeciency, color = (factor(V1))))+
  geom_line()

ggplot(total_old_new, aes(x=year))+
  geom_line(aes(y = 'effeciency'), color = "blue")+
  geom_line(aes(y = 'effeciency_total'), color = "red")


```
Filter data into only the year 2008 then apply the year closed function to ouput complete_downed as a dataframe containing all mines in 2008 and the year they closed. 

```{r test_above}
data2<- data%>%
  filter(year == 2008)
data2 <- data2%>%
  mutate(
    rank = 1:nrow(data2))


newtable <- lapply(data2$msha.id, year_closed)


newtable <- as.data.frame(do.call(rbind,newtable))

newtable <- newtable%>%
  mutate(rank = 1:nrow(newtable))
complete_downed <- full_join(data2, newtable)
complete_downed <- complete_downed%>%
  mutate(nat_trend = case_when(V1 == 2008 ~ "up",V1 == 2008 ~ "up", V1 == 2010 ~ "up", V1 != 2008 & V1 != 2010 ~ "down"))
complete_downed$appalachia <- replace_na(complete_downed$appalachia, "No")
  
```

```{r low_function}
data_copy <- data

year_bar_low <- function(yearin){
  new_data<- filter(data_copy, year == yearin)
  low_bar <-(quantile(new_data$productivity, c(0.25)))
  return(low_bar)
}
yo <- year_bar(2008)
list_low <-c()
ye <- 2008
while(ye <= 2016){
  c <- year_bar_low(ye)
  list_low<-c(list_low,c[1])
  ye= ye+1
}
```

```{r high_function}
year_bar_high <- function(yearin){
  new_data<- filter(data, year == yearin)
  high_bar <- unname(quantile(new_data$productivity, c(0.75)))
  return(high_bar)
}

list_high <-c()
ye <- 2008
while(ye <= 2016){
  b <- year_bar_high(ye)
  list_high<-c(list_high,b[1])
  ye= ye+1
}
bar_table <- as.data.frame(list_high)
bar_table<- as.data.frame(cbind(list_high, list_low))
yearer <- c(2008,2009,2010,2011,2012,2013,2014,2015,2016)
bar_table<- as.data.frame(cbind(bar_table, yearer))
```


```{r high_low}
since2008 <- data%>%
  filter(year >=2008)
mine_sum <- since2008%>%
  group_by(msha.id)%>%
  summarize(avg_prod = mean(productivity))
total_surv <- merge(x = complete_downed, y=mine_sum, by = "msha.id", all.x = TRUE)
 

total_surv<- total_surv%>%
  mutate(prod_high = case_when(avg_prod < median_productivity ~ "0", avg_prod > median_productivity ~ "1"))%>%
  select(-countystr, -operating.company.address, -operating.company)%>%
  mutate(time = V1-2007)
total_surv$appalachia <- replace_na(total_surv$appalachia, "No")
total_surv <- total_surv[order(total_surv$time),]
```

Up to this point, Total Surv isn't producing any NAs in sigfnicant catagories


```{r temp_close}
temp_close <- function(id){
  yearin = 2008
  max = 0
  switch = 0
  while(yearin <= 2016){
  current_year <- complete_open%>%
    filter(year==yearin)
  current_year <- current_year[,'msha.id']
  if ((id %in% current_year$msha.id && switch == 0) == TRUE){
    switch <-1
    max <- max +1
  }
  if ((id %in% current_year$msha.id ) == FALSE){
    if((switch == 0) == FALSE){
    switch <-0
    max <- max +1
    }
  }
  yearin <- yearin+1
  }
  if (2<max){
    return (0)
  }
  else{
    return (1)
  }
}
  
```

The first line of code here were loosing lots of data entries to merge and I'm not sure why. It looks like full_join fixed this, but I'm not entirely sure.

A key thing moving forward is that alive is defined here as if the mine is in 2016. Will need to change that for the other runs.

This full join causes all the NAs. Total_surv is only 1203. I'm wondering if because it's full join it screws up when something is in one but not the other. There are entries with appalacia as both NA
```{r remake_total}
grand <- right_join(x = since2008, y=total_surv, by = "msha.id")

grand<- grand%>%
  mutate(surface = case_when(mine.type.x != "Surface" ~ 0, mine.type.x == "Surface" ~ 1))%>%
  mutate(union_code = case_when(
    is.na(union.code.x) ~ 0, !is.na(union.code.x) ~ 1
  ))
grand<- merge(x= grand, y = bar_table, by.x = "year.x", by.y = "yearer")
grand<- grand%>% 
  mutate(high = case_when(
    productivity.x > list_high ~ 1, productivity.x <= list_high ~ 0))%>%
   mutate(alive = case_when(V1 == 2016 ~ 0, V1 != 2016 ~ 1))%>%
  mutate(low = case_when(
    productivity.x < list_low ~ 1, productivity.x >= list_low ~ 0))%>%
  mutate(time = year.x - 2007)%>%
  rename(appalachia = appalachia.y)
```

Apply the function to tell if a mine underwent a temporary closure

Having trouble here with getting double the variables I should be

```{r apply_test}
test_table <- lapply(grand$msha.id, temp_close)

new_test_table <- do.call(rbind.data.frame, test_table)

colnames(new_test_table) <- "temporary_close"
new_test_table <- new_test_table%>%
  mutate(rank = 1:nrow(new_test_table))


grand_with_test <- left_join(grand, new_test_table, by = "rank")
head(grand_with_test)
head(new_test_table)
```

Select only variables desired from the Grand datas frame. Also set the frame for grand

```{r clean_grand}
clean_grand <- grand %>%
  select(year.x, msha.id, mine.name.x, V1, alive, avg_prod, appalachia, productivity.x, surface, union_code, nat_trend)%>%
  mutate(time = year.x - 2007)
```

Here I need to figure out how to identify which out of the not alive mines. I believe I need to paste the selection and mutate from the clean grand but this could be tweaked

```{r never_close}
never_close <- grand_with_test%>%
  mutate(alive = case_when(alive == 1 ~ temporary_close, alive != 1 ~ alive))%>%
  mutate(time = year.x - 2007)
  
```

Create a data set to run survivial metrics that excludes any mines that temporarily closed

```{r removed}
removed<- grand_with_test%>%
  filter(temporary_close != 0)%>%
  mutate(time = year.x - 2007)
```


```{r univariate_cox}
surv_object<- Surv(time = edited_surv$time, event = edited_surv$alive)

surv_object2<- Surv(time = edited_surv2$time, event = edited_surv2$alive)
```



Need to decipher if I want grand or clean grand (or equivelents w new dataset)

Also why are some missing? Different numbers missing in different data sets. Bad! "deleted due to missing"

Looks like issue is they all have different numbers to start with

```{r surv_fit}
surv_diff <- survdiff(Surv(time, alive) ~ appalachia, data = clean_grand)
surv_diff

surv_diff_removed <- survdiff(Surv(time,alive)~ appalachia, data = removed)
surv_diff_removed

surv_diff_never_close <- survdiff(Surv(time,alive)~appalachia, data = never_close)
surv_diff_never_close
```

```{r surv_fit_high_low}
surv_diff <- survdiff(Surv(time, alive) ~ low, data = grand)

fithl<- survfit(Surv(time, alive) ~ low, data = grand)
summary(fithl)
fithl$surv
fithlp <- ggsurvplot(fithl, data = grand, pval = TRUE)
fithlp

surv_diff_removed <- survdiff(Surv(time, alive) ~ low, data = removed)

fithl_removed<- survfit(Surv(time, alive) ~ low, data = removed)
summary(fithl_removed)
fithlp_removed <- ggsurvplot(fithl_removed, data = removed, pval = TRUE)
fithlp_removed

surv_diff_never_close <- survdiff(Surv(time, alive) ~ low, data = never_close)

fithl_never_close<- survfit(Surv(time, alive) ~ low, data = never_close)
summary(fithl_never_close)

fithlp_never_close <- ggsurvplot(fithl_never_close, data = removed, pval = TRUE)
fithlp_never_close
```

```{r survivial_fit}
fit1<- survfit(Surv(time, alive) ~ appalachia, data = grand)
summary(fit1)
fit1 <- ggsurvplot(fit1, data = grand, pval = TRUE)
fit1

fit1_removed<- survfit(Surv(time, alive) ~ appalachia, data = removed)
summary(fit1_removed)
fit1_removed <- ggsurvplot(fit1_removed, data = removed, pval = TRUE)
fit1_removed

fit1_never_close<- survfit(Surv(time, alive) ~ appalachia, data = never_close)
summary(fit1_never_close)
fit1_never_close <- ggsurvplot(fit1_never_close, data = never_close, pval = TRUE)
fit1_never_close
```



```{r high_and_low}
datahl<- grand%>%
  filter(low == 1 | high == 1)%>%
  mutate(highest = case_when(low == 1 ~ 0, high == 1 ~ 1))
fithl<- survfit(Surv(time, alive) ~ highest, data = datahl)
summary(fithl)

fithlg <- ggsurvplot(fithl, data = datahl, pval = TRUE)
fithlg

datahl_removed<- removed%>%
  filter(low == 1 | high == 1)%>%
  mutate(highest = case_when(low == 1 ~ 0, high == 1 ~ 1))
fithl_removed<- survfit(Surv(time, alive) ~ highest, data = datahl_removed)
summary(fithl_removed)

fithlg_removed <- ggsurvplot(fithl_removed, data = datahl_removed, pval = TRUE)
fithlg_removed

datahl_never_close<- never_close%>%
  filter(low == 1 | high == 1)%>%
  mutate(highest = case_when(low == 1 ~ 0, high == 1 ~ 1))
fithl_never_close<- survfit(Surv(time, alive) ~ highest, data = datahl_never_close)
summary(fithl_never_close)

fithlg_never_close <- ggsurvplot(fithl_never_close, data = datahl_never_close, pval = TRUE)
fithlg_never_close
```

```{r national_trend}
fit5<- survfit(Surv(time, alive) ~ nat_trend, data = clean_grand)
summary(fit5)

fito <- ggsurvplot(fit5, data = clean_grand, pval = TRUE)
fito

fit5_removed<- survfit(Surv(time, alive) ~ nat_trend, data = removed)
summary(fit5_removed)

fito_removed <- ggsurvplot(fit5_removed, data = removed, pval = TRUE)
fito_removed

fit5_never_close<- survfit(Surv(time, alive) ~ nat_trend, data = never_close)
summary(fit5_never_close)

fito_never_close <- ggsurvplot(fit5_never_close, data = never_close, pval = TRUE)
fito_never_close
```

```{r survivial_fit_union}
fit2<- survfit(Surv(time, alive) ~ union_code, data = clean_grand)

fitm <- ggsurvplot(fit2, data = clean_grand, pval = TRUE)
fitm
summary(fit2)

fit2_removed<- survfit(Surv(time, alive) ~ union_code, data = removed)

summary(fit2_removed)

fitm_removed <- ggsurvplot(fit2_removed, data = removed, pval = TRUE)
fitm_removed

fit2_never_close<- survfit(Surv(time, alive) ~ union_code, data = never_close)

summary(fit2_never_close)

fitm_never_close <- ggsurvplot(fit2_never_close, data = never_close, pval = TRUE)
fitm_never_close
```

```{r survivial_fit_surface}
fit3<- survfit(Surv(time, alive) ~ surface, data = clean_grand)

fitn <- ggsurvplot(fit3, data = clean_grand, pval = TRUE)
summary(fit3)
fitn

fit3_removed<- survfit(Surv(time, alive) ~ surface, data = removed)
summary(fit3_removed)
fitn_removed <- ggsurvplot(fit3_removed, data = removed, pval = TRUE)
fitn_removed

fit3_never_close<- survfit(Surv(time, alive) ~ surface, data = never_close)
summary(fit3_never_close)
fitn_never_close <- ggsurvplot(fit3_never_close, data = never_close, pval = TRUE)
summary(fitn_never_close)
fitn_never_close
```

```{r cox_haz}
fit_cox<- coxph(Surv(time, alive) ~  appalachia + union_code + surface, data = grand)
summary(fit_cox)
vif(fit_cox)
fit_cox_removed<- coxph(Surv(time, alive) ~  appalachia + union_code + surface, data = removed)
summary(fit_cox_removed)
vif(fit_cox_removed)

fit_cox_never_close<- coxph(Surv(time, alive) ~  appalachia + union_code + surface, data = never_close)
summary(fit_cox_never_close)
vif(fit_cox_never_close)

```

```{r univariate_analysis}

fit_cox_uni_union<- coxph(Surv(time, alive) ~ union_code, data = grand)
summary(fit_cox_uni_union)

fit_cox_uni_low<- coxph(Surv(time, alive) ~ low, data = grand)
summary(fit_cox_uni_low)

fit_cox_uni_surface <- coxph(Surv(time, alive) ~ surface, data = grand)
summary(fit_cox_uni_surface)

fit_cox_uni_high<- coxph(Surv(time, alive) ~ high, data = grand)
summary(fit_cox_uni_high)


fit_cox_uni_union_removed<- coxph(Surv(time, alive) ~ union_code, data = removed)
summary(fit_cox_uni_union_removed)

fit_cox_uni_low_removed<- coxph(Surv(time, alive) ~ low, data = removed)
summary(fit_cox_uni_low_removed)

fit_cox_uni_surface_removed <- coxph(Surv(time, alive) ~ surface, data = removed)
summary(fit_cox_uni_surface_removed)

fit_cox_uni_high_removed<- coxph(Surv(time, alive) ~ high, data = removed)
summary(fit_cox_uni_high_removed)

fit_cox_uni_union_never_close<- coxph(Surv(time, alive) ~ union_code, data = never_close)
summary(fit_cox_uni_union_never_close)

fit_cox_uni_low_never_close<- coxph(Surv(time, alive) ~ low, data = never_close)
summary(fit_cox_uni_low_never_close)

fit_cox_uni_surface_never_close <- coxph(Surv(time, alive) ~ surface, data = never_close)
summary(fit_cox_uni_surface_never_close)

fit_cox_uni_high_never_close<- coxph(Surv(time, alive) ~ high, data = never_close)
summary(fit_cox_uni_high_never_close)





to_vif <- coxph(Surv(time, alive) ~ appalachia +union_code + surface, data = grand)

is_vif <-vif(lm(alive ~ appalachia + union_code +, data=grand))
vif(to_vif)


```



```{r plot_cox}
forest <- ggforest(coxph(Surv(time, alive) ~ appalachia +union_code + low, data = grand))
forest

forest_removed <- ggforest(coxph(Surv(time, alive) ~ appalachia +union_code + low, data = removed))
forest_removed

forest_never_close <- ggforest(coxph(Surv(time, alive) ~ appalachia +union_code + low, data = never_close))
forest_never_close
```

```{r surv_expo}
fit_cox_uni<- coxph(Surv(time, alive) ~ nat_trend, data = edited_surv)
summary(fit_cox_uni)
```
